---
title: "US NOOA Analysis"
author: "Moses Otieno"
date: "August 10, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Libraries,include=F}
#--- Libraries
library(kableExtra)
library(gt)
library(gtsummary)
library(magrittr)
library(lubridate)
library(tidyverse)
library(table1)
library(modelr)
library(moderndive)
library(naniar)
```



```{r importdata, cache=TRUE}

stormdata <- read.csv("./data/rawdata/repdata_data_StormData.csv.bz2", header = TRUE, sep = ",", quote = "\"")
```

### Data Preprocessing

```{r data_preparation}
#---- Have a look at raw data

stormdata <- as_tibble(stormdata)   # Convert to tibbles to easen work
stormdata

#--- Change the column names to lower

names(stormdata) <- str_to_lower(names(stormdata))


dim(stormdata)                     # Dimensions of the dataset
head(stormdata)                    # Have a look at first six rows
tail(stormdata)                    # Have a look at the last six rows

sample_n(stormdata, 10)            # Sample 10 observations
glimpse(stormdata)                 # Structure of the dataset
summary(stormdata)                 # General summary of the whole dataset

#---- Variable transformation

#-- Dates and times 

stormdata %>% 
  select(matches("[Dd]ate"), contains("date"))


stormdata$bgn_date <-   as.Date(stormdata$bgn_date, "%m/%d/%Y %H:%M:%S")
stormdata



#--- Missing values 
miss_var_table(stormdata)

miss_var_summary(stormdata)

stormdata %>% 
  miss_var_summary %>% 
  filter( pct_miss>= 90)

stormdata %>% 
  add_prop_miss() %>% 
  arrange(desc(prop_miss_all)) %>% 
  mutate(prop_miss_all = round(prop_miss_all * 100), 1) %>% 
  select(prop_miss_all)


gg_miss_var(stormdata) + labs(y = "Look at all the missing ones") + theme_bw() + theme(
  axis.title.y = element_text(size = "8")
)

#-- Drop the single variable with all missing observations
  
stormdata <- select(stormdata, -countyendn)


#-- Check whether there are duplicates in the dataset
sum(duplicated(stormdata))

```

Clearly this dataset is is not largely affected by the missing values. About 90% of the variables have no missing. To add on that the highest missing observed in a given observation was 11%. This again indicates that the data collection was well done This definitely indicates high level of completeness of data. 

We have `r sum(duplicated(stormdata))` duplicates in the entire dataset.

